---
title: "SPACIC manuscript"
author: "JacquelineHoppenreijs"
date: "`r Sys.Date()`"
bibliography: references.bib
output: word_document
---

```{r setting the scene, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

install.packages("pacman")
library(pacman)
p_load(readxl, tidyverse, plyr, dplyr, lattice, ggplot2, png, colorblindcheck, broom, vegan, patchwork, scales, afex, effectsize, ltm)

# Check whether the used colours are colourblindness-friendly
palette <- c("#104e8b", "#458b00", "#eead0e", "#9F2B68") 
palette_check(palette)
palette_bivariate_plot(palette)

#Check which packages and versions are included
package_versions <- function(packages=NULL, base=FALSE, sort=FALSE) {
    if (is.null(packages))
        packages <- names(sessionInfo()$otherPkgs)
    if (base)
        packages <- c("base", packages)
    if (sort)
        packages <- sort(packages)
    if (!is.null(packages)) {
        tmp <- data.frame(sapply(sapply(packages, packageVersion, simplify=FALSE), as.character))
        colnames(tmp) <- "version"
        data.frame(package=rownames(tmp), version=tmp[1], row.names=NULL)
    }
}
```

# File description
This Rmarkdown document can be used to analyse plant and insect data, and is tailored for use in the FORMAS project "Lära och Leva med Insekter", led by Peter Lampert (Karlstad University). He can be contacted at peter.lampert@kau.se. The document is written by Jacqueline Hoppenreijs (jacquelinehthoppenreijs@gmail.com).

# Software specifications
Built with R version `r getRversion()`. 

Used packages + versions: `package_versions()`.


```{r loading data, include = FALSE}
schooldata <- read_excel("./SPACIC/FullData_JH20241120.xlsx", sheet = "RepeatedMeasures")
schooldata$School <- as.factor(schooldata$School)
schooldata$Treatment <- as.factor(schooldata$Treatment)

str(schooldata)

control_data <- subset(schooldata, Treatment == "Control")
intervention_data <- subset(schooldata, Treatment == "Intervention")

schooldata$School_ordered = factor(schooldata$School, levels=c("NC", "ST","IC", "JM", "OZ"))
```

# Likert scale distributions
## SPACIC

```{r visualising full SPACIC, include = FALSE}
SPACIC_B_plot <- ggplot(schooldata, aes(x = Before_Full_SPACIP_mean, fill = Treatment)) +
  facet_grid( School_ordered ~.) +
  geom_histogram(binwidth = 0.2) + 
  scale_x_continuous("Mean SPACIC before") +
  scale_y_continuous("Number of pupils", limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) +
  theme_classic() +
  theme(strip.text = element_blank()) +
  guides(fill = FALSE)  

SPACIC_A_plot <- ggplot(schooldata, aes(x = After_Full_SPACIP_mean, fill = Treatment)) +
  facet_grid( School_ordered ~.) +
  geom_histogram(binwidth = 0.2) + 
  scale_x_continuous("Mean SPACIC after") +
  scale_y_continuous("", limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) +
  theme_classic()

SPACIC_plots <- SPACIC_B_plot + SPACIC_A_plot + plot_layout(nrow = 1)

ggsave(filename = "SPACIC_plots_20241122.png", 
       plot = SPACIC_plots, path = "SPACIC")
```

```{r full SPACIC full picture, include = TRUE}
SPACIC_plots
```

## K nowledge
```{r visualising full K, include = FALSE}
Kfull_B_plot <- ggplot(schooldata, aes(x = Before_K_mean, fill = Treatment)) +
  facet_grid( School_ordered ~.) +
  geom_histogram(binwidth = 0.2) + 
  scale_x_continuous("Mean K before") +
  scale_y_continuous("Number of pupils", limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) +
  theme_classic() +
  theme(strip.text = element_blank()) +
  guides(fill = FALSE)  

Kfull_A_plot <- ggplot(schooldata, aes(x = After_K_mean, fill = Treatment)) +
  facet_grid( School_ordered ~.) +
  geom_histogram(binwidth = 0.2) + 
  scale_x_continuous("Mean K after") +
  scale_y_continuous("", limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) +
  theme_classic()

Kfull_plots <- Kfull_B_plot + Kfull_A_plot + plot_layout(nrow = 1)

ggsave(filename = "Kfull_plots_20241122.png", 
       plot = Kfull_plots, path = "SPACIC")
```

```{r full K full picture, include = TRUE}
Kfull_plots
```

## C onfidence
```{r visualising full C, include = FALSE}
Cfull_B_plot <- ggplot(schooldata, aes(x = Before_C_mean, fill = Treatment)) +
  facet_grid( School_ordered ~.) +
  geom_histogram(binwidth = 0.2) + 
  scale_x_continuous("Mean C before") +
  scale_y_continuous("Number of pupils", limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) +
  theme_classic() +
  theme(strip.text = element_blank()) +
  guides(fill = FALSE)  

Cfull_A_plot <- ggplot(schooldata, aes(x = After_C_mean, fill = Treatment)) +
  facet_grid( School_ordered ~.) +
  geom_histogram(binwidth = 0.2) + 
  scale_x_continuous("Mean C after") +
  scale_y_continuous("", limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) +
  theme_classic()

Cfull_plots <- Cfull_B_plot + Cfull_A_plot + plot_layout(nrow = 1)

ggsave(filename = "Cfull_plots_20241122.png", 
       plot = Cfull_plots, path = "SPACIC")
```

```{r full C full picture, include = TRUE}
Cfull_plots
```

## W illingness
```{r visualising full C, include = FALSE}
Wfull_B_plot <- ggplot(schooldata, aes(x = Before_W_mean, fill = Treatment)) +
  facet_grid(School_ordered ~.) +
  geom_histogram(binwidth = 0.2) + 
  scale_x_continuous("Mean W before") +
  scale_y_continuous("Number of pupils", limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) +
  theme_classic() +
  theme(strip.text = element_blank()) +
  guides(fill = FALSE)  

Wfull_A_plot <- ggplot(schooldata, aes(x = After_W_mean, fill = Treatment)) +
  facet_grid(School_ordered ~.) +
  geom_histogram(binwidth = 0.2) + 
  scale_x_continuous("Mean W after") +
  scale_y_continuous("", limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) +
  theme_classic()

Wfull_plots <- Wfull_B_plot + Wfull_A_plot + plot_layout(nrow = 1)

ggsave(filename = "Wfull_plots_20241122.png", 
       plot = Wfull_plots, path = "SPACIC")
```

```{r full C full picture, include = TRUE}
Wfull_plots
```

# SPACIC, K, C and W scores
## SPACIC

```{r loading more data, include = FALSE}
scoredata <- read_excel("./SPACIC/FullData_JH20241122.xlsx", sheet = "SubsetTidy")
scoredata$School <- as.factor(scoredata$School)
scoredata$Treatment <- as.factor(scoredata$Treatment)
scoredata$Timing <- as.factor(scoredata$Timing)

str(scoredata)

scoredata_b <- subset(scoredata, Timing == "Pre-Test")
scoredata_a <- subset(scoredata, Timing == "Post-Test")

control_data_b <- subset(scoredata_b, Treatment == "Control")
intervention_data_b <- subset(scoredata_b, Treatment == "Intervention")

control_data_a <- subset(scoredata_a, Treatment == "Control")
intervention_data_a <- subset(scoredata_a, Treatment == "Intervention")

scoredata$Timing_ordered = factor(scoredata$Timing, levels=c("Pre-Test", "Post-Test"))
```

```{r visualising mean SPACIC, include = FALSE}
meanSPACIC_b_Control <- mean(control_data_b$Full_SPACIP_mean)
seSPACIC_b_Control <- sd(control_data_b$Full_SPACIP_mean)/sqrt(length(control_data_b$Full_SPACIP_mean))

meanSPACIC_b_Intervention <- mean(intervention_data_b$Full_SPACIP_mean)
seSPACIC_b_Intervention <- sd(intervention_data_b$Full_SPACIP_mean)/sqrt(length(intervention_data_b$Full_SPACIP_mean))

meanSPACIC_a_Control <- mean(control_data_a$Full_SPACIP_mean)
seSPACIC_a_Control <- sd(control_data_a$Full_SPACIP_mean)/sqrt(length(control_data_a$Full_SPACIP_mean))

meanSPACIC_a_Intervention <- mean(intervention_data_a$Full_SPACIP_mean)
seSPACIC_a_Intervention <- sd(intervention_data_a$Full_SPACIP_mean)/sqrt(length(intervention_data_a$Full_SPACIP_mean))

SPACIC_data <- data.frame(
  Group = rep(c("Control", "Intervention"), each = 2),
  Moment = rep(c("Pre-Test", "Post-Test"), times = 2),
  Mean = c(meanSPACIC_b_Control, meanSPACIC_a_Control, 
           meanSPACIC_b_Intervention, meanSPACIC_a_Intervention),
  SE = c(seSPACIC_b_Control, seSPACIC_a_Control, 
         seSPACIC_b_Intervention, seSPACIC_a_Intervention)
)

SPACIC_data$Moment = factor(SPACIC_data$Moment, levels=c("Pre-Test", "Post-Test"))
SPACIC_data$Group <- factor(SPACIC_data$Group, levels = c("Control", "Intervention"))

SPACIC_plot <- ggplot(SPACIC_data, aes(x = Moment, y = Mean, group = Group)) +
  geom_line(aes(color = Group), linewidth = 1.2) + 
  geom_point(aes(color = Group), size = 4) + 
  geom_errorbar(aes(ymin = Mean - SE, ymax = Mean + SE, color = Group), width = 0.2, size = 1.2) + 
  scale_color_manual(values = c("Control" = "azure4", "Intervention" = "black")) +
  scale_x_discrete(name = "") +
  scale_y_continuous(name = "Mean SPACIC ± SE", limits = c(1, 5)) +
  theme_classic(base_size = 24) + 
  theme(axis.title = element_text(size = 24, face = "bold"), 
        axis.text.x = element_blank(), axis.ticks.x = element_blank(), 
        axis.text = element_text(size = 24)) +
  theme(legend.position = c(0.05, 0.95), 
    legend.justification = c("left", "top"),
    legend.title = element_blank(), 
    legend.background = element_rect(fill = "white", color = "black", linewidth = 0.5)) 
SPACIC_plot
```

## K nowledge
```{r visualising mean K, include = FALSE}
meanKfull_b_Control <- mean(control_data_b$K_mean)
seKfull_b_Control <- sd(control_data_b$K_mean)/sqrt(length(control_data_b$K_mean))

meanKfull_b_Intervention <- mean(intervention_data_b$K_mean)
seKfull_b_Intervention <- sd(intervention_data_b$K_mean)/sqrt(length(intervention_data_b$K_mean))

meanKfull_a_Control <- mean(control_data_a$K_mean)
seKfull_a_Control <- sd(control_data_a$K_mean)/sqrt(length(control_data_a$K_mean))

meanKfull_a_Intervention <- mean(intervention_data_a$K_mean)
seKfull_a_Intervention <- sd(intervention_data_a$K_mean)/sqrt(length(intervention_data_a$K_mean))

Kfull_data <- data.frame(
  Group = rep(c("Control", "Intervention"), each = 2),
  Moment = rep(c("Pre-Test", "Post-Test"), times = 2),
  Mean = c(meanKfull_b_Control, meanKfull_a_Control, 
           meanKfull_b_Intervention, meanKfull_a_Intervention),
  SE = c(seKfull_b_Control, seKfull_a_Control, 
         seKfull_b_Intervention, seKfull_a_Intervention)
)

Kfull_data$Moment = factor(Kfull_data$Moment, levels=c("Pre-Test", "Post-Test"))
Kfull_data$Group <- factor(Kfull_data$Group, levels = c("Control", "Intervention"))

Kfull_plot <- ggplot(Kfull_data, aes(x = Moment, y = Mean, group = Group)) +
  geom_line(aes(color = Group), linewidth = 1.2) + 
  geom_point(aes(color = Group), size = 4) + 
  geom_errorbar(aes(ymin = Mean - SE, ymax = Mean + SE, color = Group), width = 0.2, size = 1.2) + 
  scale_color_manual(values = c("Control" = "azure4", "Intervention" = "black")) +
  scale_x_discrete(name = "", labels = c("Pre-test", "Post-test")) +
  scale_y_continuous(name = "Mean K ± SE", limits = c(1, 5)) +
  theme_classic(base_size = 24) + 
  theme(axis.title = element_text(size = 24, face = "bold"), axis.text = element_text(size = 24)) +
  theme(legend.position = "none") #+
  #theme(legend.position = c(0.05, 0.95), 
    #legend.justification = c("left", "top"),
    #legend.title = element_blank(), 
    #legend.background = element_rect(fill = "white", color = "black", linewidth = 0.5)) 
Kfull_plot
```

## C onfidence
```{r visualising mean C, include = FALSE}
meanCfull_b_Control <- mean(control_data_b$C_mean)
seCfull_b_Control <- sd(control_data_b$C_mean)/sqrt(length(control_data_b$C_mean))

meanCfull_b_Intervention <- mean(intervention_data_b$C_mean)
seCfull_b_Intervention <- sd(intervention_data_b$C_mean)/sqrt(length(intervention_data_b$C_mean))

meanCfull_a_Control <- mean(control_data_a$C_mean)
seCfull_a_Control <- sd(control_data_a$C_mean)/sqrt(length(control_data_a$C_mean))

meanCfull_a_Intervention <- mean(intervention_data_a$C_mean)
seCfull_a_Intervention <- sd(intervention_data_a$C_mean)/sqrt(length(intervention_data_a$C_mean))

Cfull_data <- data.frame(
  Group = rep(c("Control", "Intervention"), each = 2),
  Moment = rep(c("Pre-Test", "Post-Test"), times = 2),
  Mean = c(meanCfull_b_Control, meanCfull_a_Control, 
           meanCfull_b_Intervention, meanCfull_a_Intervention),
  SE = c(seCfull_b_Control, seCfull_a_Control, 
         seCfull_b_Intervention, seCfull_a_Intervention)
)

Cfull_data$Moment = factor(Cfull_data$Moment, levels=c("Pre-Test", "Post-Test"))
Cfull_data$Group <- factor(Cfull_data$Group, levels = c("Control", "Intervention"))

Cfull_plot <- ggplot(Cfull_data, aes(x = Moment, y = Mean, group = Group)) +
  geom_line(aes(color = Group), linewidth = 1.2) + 
  geom_point(aes(color = Group), size = 4) + 
  geom_errorbar(aes(ymin = Mean - SE, ymax = Mean + SE, color = Group), width = 0.2, size = 1.2) + 
  scale_color_manual(values = c("Control" = "azure4", "Intervention" = "black")) +
  scale_x_discrete(name = "") +
  scale_y_continuous(name = "Mean C ± SE", limits = c(1, 5)) +
  theme_classic(base_size = 24) + 
  theme(axis.title = element_text(size = 24, face = "bold"), 
        axis.text.x = element_blank(), axis.ticks.x = element_blank(), 
        axis.text = element_text(size = 24)) +
  theme(legend.position = "none") #+
  #theme(legend.position = c(0.05, 0.95), 
    #legend.justification = c("left", "top"),
    #legend.title = element_blank(), 
    #legend.background = element_rect(fill = "white", color = "black", linewidth = 0.5)) 
Cfull_plot
```

## W illingness
```{r visualising mean W, include = FALSE}
meanWfull_b_Control <- mean(control_data_b$W_mean)
seWfull_b_Control <- sd(control_data_b$W_mean)/sqrt(length(control_data_b$W_mean))

meanWfull_b_Intervention <- mean(intervention_data_b$W_mean)
seWfull_b_Intervention <- sd(intervention_data_b$W_mean)/sqrt(length(intervention_data_b$W_mean))

meanWfull_a_Control <- mean(control_data_a$W_mean)
seWfull_a_Control <- sd(control_data_a$W_mean)/sqrt(length(control_data_a$W_mean))

meanWfull_a_Intervention <- mean(intervention_data_a$W_mean)
seWfull_a_Intervention <- sd(intervention_data_a$W_mean)/sqrt(length(intervention_data_a$W_mean))

Wfull_data <- data.frame(
  Group = rep(c("Control", "Intervention"), each = 2),
  Moment = rep(c("Pre-Test", "Post-Test"), times = 2),
  Mean = c(meanWfull_b_Control, meanWfull_a_Control, 
           meanWfull_b_Intervention, meanWfull_a_Intervention),
  SE = c(seWfull_b_Control, seWfull_a_Control, 
         seWfull_b_Intervention, seWfull_a_Intervention)
)

Wfull_data$Moment = factor(Wfull_data$Moment, levels=c("Pre-Test", "Post-Test"))
Wfull_data$Group <- factor(Wfull_data$Group, levels = c("Control", "Intervention"))

Wfull_plot <- ggplot(Wfull_data, aes(x = Moment, y = Mean, group = Group)) +
  geom_line(aes(color = Group), linewidth = 1.2) + 
  geom_point(aes(color = Group), size = 4) + 
  geom_errorbar(aes(ymin = Mean - SE, ymax = Mean + SE, color = Group), width = 0.2, size = 1.2) + 
  scale_color_manual(values = c("Control" = "azure4", "Intervention" = "black")) +
  scale_x_discrete(name = "", labels = c("Pre-test", "Post-test")) +
  scale_y_continuous(name = "Mean W ± SE", limits = c(1, 5)) +
  theme_classic(base_size = 24) + 
  theme(axis.title = element_text(size = 24, face = "bold"), 
        axis.text = element_text(size = 24)) +
  theme(legend.position = "none") #+
  #theme(legend.position = c(0.05, 0.95), 
    #legend.justification = c("left", "top"),
    #legend.title = element_blank(), 
    #legend.background = element_rect(fill = "white", color = "black", linewidth = 0.5)) 
Wfull_plot

```

```{r combine all line plots and save, include = TRUE}
Line_plots_20241128 <- SPACIC_plot + Cfull_plot + Kfull_plot + Wfull_plot + plot_layout(nrow = 2)
Line_plots_20241128

ggsave(filename = "Line_plots_20241128.png", 
       plot = Line_plots_20241128, path = "SPACIC")
```

# ANOVA
```{r ANOVA, include = TRUE}
str(scoredata)
scoredata$Code <- as.factor(scoredata$Code)

#Between subjects: Treatment
#Within subjects: Timing

model_scoreSPACIC <- aov_car(Full_SPACIP_mean ~ Timing * Treatment + Error(Code/Timing), data = scoredata)
summary(model_scoreSPACIC)
eta_squared(model_scoreSPACIC, ci = 0.90)
eta_squared(model_scoreSPACIC, generalized = "Treatment")

model_scoreKfull <- aov_car(K_mean ~ Timing * Treatment + Error(Code/Timing), data = scoredata)
summary(model_scoreKfull)
eta_squared(model_scoreKfull, ci = 0.90)
eta_squared(model_scoreKfull, generalized = "Treatment")

model_scoreKdir <- aov_car(K_mean_dir ~ Timing * Treatment + Error(Code/Timing), data = scoredata)
summary(model_scoreKdir)
eta_squared(model_scoreKdir, ci = 0.90)
eta_squared(model_scoreKdir, generalized = "Treatment")

model_scoreKindir <- aov_car(K_mean_ind ~ Timing * Treatment + Error(Code/Timing), data = scoredata)
summary(model_scoreKindir)
eta_squared(model_scoreKindir, ci = 0.90)
eta_squared(model_scoreKindir, generalized = "Treatment")

model_scoreCfull <- aov_car(C_mean ~ Timing * Treatment + Error(Code/Timing), data = scoredata)
summary(model_scoreCfull)
eta_squared(model_scoreCfull, ci = 0.90)
eta_squared(model_scoreCfull, generalized = "Treatment")

model_scoreCdir <- aov_car(C_mean_dir ~ Timing * Treatment + Error(Code/Timing), data = scoredata)
summary(model_scoreCdir)
eta_squared(model_scoreCdir, ci = 0.90)
eta_squared(model_scoreCdir, generalized = "Treatment")

model_scoreCindir <- aov_car(C_mean_ind ~ Timing * Treatment + Error(Code/Timing), data = scoredata)
summary(model_scoreCindir)
eta_squared(model_scoreCindir, ci = 0.90)
eta_squared(model_scoreCindir, generalized = "Treatment")

model_scoreWfull <- aov_car(W_mean ~ Timing * Treatment + Error(Code/Timing), data = scoredata)
summary(model_scoreWfull)
eta_squared(model_scoreWfull, ci = 0.90)
eta_squared(model_scoreWfull, generalized = "Treatment")

model_scoreWdir <- aov_car(W_mean_dir ~ Timing * Treatment + Error(Code/Timing), data = scoredata)
summary(model_scoreWdir)
eta_squared(model_scoreWdir, ci = 0.90)
eta_squared(model_scoreWdir, generalized = "Treatment")

model_scoreWindir <- aov_car(W_mean_ind ~ Timing * Treatment + Error(Code/Timing), data = scoredata)
summary(model_scoreWindir)
eta_squared(model_scoreWindir, ci = 0.90)
eta_squared(model_scoreWindir, generalized = "Treatment")

```

# Descriptive stats

## Table 1

### Full dataset
```{r loading descriptive data full, include = FALSE}
descr_data_full <- read_excel("./SPACIC/FullData_JH20241125.xlsx", sheet = "MergeFileTidy")
descr_data_full$School <- as.factor(descr_data_full$School)
descr_data_full$Treatment <- as.factor(descr_data_full$Treatment)
descr_data_full$Timing <- as.factor(descr_data_full$Timing)
descr_data_full$Gender <- as.factor(descr_data_full$Gender)

str(descr_data_full)

descr_data_full$Treatment = factor(descr_data_full$Treatment, levels=c("Intervention", "Control"))
descr_data_full$Timing = factor(descr_data_full$Timing, levels=c("Before", "During", "After"))
```

```{r build df table 1a, include = FALSE}
# Recalculating age
descr_data_full$Age <- ifelse(descr_data_full$Age == 3, 12,
                              ifelse(descr_data_full$Age == 4, 13,
                                     ifelse(descr_data_full$Age == 5, 14,
                                            ifelse(descr_data_full$Age == 6, 15, descr_data_full$Age))))


gender_full_df <- ddply(descr_data_full, ~ Treatment*Timing*Gender, summarise, 
                       n_pupils = length(Full_SPACIP_mean), 
                       MeanAge = mean(Age))
gender_full_df
```

### RM ANOVA subset
```{r loading descriptive data subset, include = FALSE}
descr_data_subs <- read_excel("./SPACIC/FullData_JH20241125.xlsx", sheet = "SubsetTidy")
descr_data_subs$School <- as.factor(descr_data_subs$School)
descr_data_subs$Treatment <- as.factor(descr_data_subs$Treatment)
descr_data_subs$Timing <- as.factor(descr_data_subs$Timing)
descr_data_subs$Gender <- as.factor(descr_data_subs$Gender)

str(descr_data_subs)

descr_data_subs$Treatment = factor(descr_data_subs$Treatment, levels=c("Intervention", "Control"))
descr_data_subs$Timing = factor(descr_data_subs$Timing, levels=c("Pre-Test", "Post-Test"))
```

```{r build df table 1b, include = FALSE}
# Recalculating age
descr_data_subs$Age <- ifelse(descr_data_subs$Age == 3, 12,
                              ifelse(descr_data_subs$Age == 4, 13,
                                     ifelse(descr_data_subs$Age == 5, 14,
                                            ifelse(descr_data_subs$Age == 6, 15, descr_data_subs$Age))))

gender_subs_df <- ddply(descr_data_subs, ~ Treatment*Timing*Gender, summarise, 
                       n_pupils = length(Full_SPACIP_mean), 
                       MeanAge = mean(Age))
gender_subs_df
```

## Table 2

### Full dataset
```{r build df table 2a, include = FALSE}
descr_full_df2a <- ddply(descr_data_full, ~ Treatment*Timing, summarise, 
                       n_pupils = length(Full_SPACIP_mean), 
                       MeanSPACIC = mean(Full_SPACIP_mean), 
                       SESPACIC = sd(Full_SPACIP_mean)/sqrt(length(Full_SPACIP_mean)), 
                       MeanKfull = mean(K_mean), 
                       SEKfull = sd(K_mean)/sqrt(length(K_mean)),
                       MeanK_mean_dir = mean(K_mean_dir), 
                       SEK_mean_dir = sd(K_mean_dir)/sqrt(length(K_mean_dir)),
                       MeanK_mean_ind = mean(K_mean_ind), 
                       SEK_mean_ind = sd(K_mean_ind)/sqrt(length(K_mean_ind)), 
                       MeanCfull = mean(C_mean), 
                       SECfull = sd(C_mean)/sqrt(length(C_mean)),
                       MeanC_mean_dir = mean(C_mean_dir), 
                       SEC_mean_dir = sd(C_mean_dir)/sqrt(length(C_mean_dir)),
                       MeanC_mean_ind = mean(C_mean_ind), 
                       SEC_mean_ind = sd(C_mean_ind)/sqrt(length(C_mean_ind)), 
                       MeanWfull = mean(W_mean), 
                       SEWfull = sd(W_mean)/sqrt(length(W_mean)),
                       MeanW_mean_dir = mean(W_mean_dir), 
                       SEW_mean_dir = sd(W_mean_dir)/sqrt(length(W_mean_dir)),
                       MeanW_mean_ind = mean(W_mean_ind), 
                       SEW_mean_ind = sd(W_mean_ind)/sqrt(length(W_mean_ind)))
descr_full_df2a
```

#### Cronbach's alpha
```{r calculate for the full dataset, include = FALSE}
str(descr_data_full)

descr_data_full_before <- subset(descr_data_full, Timing == "Before")
str(descr_data_full_before)
subs <- descr_data_full_before[, 28:35]
str(subs)

SPACIC_items <- descr_data_full_before[, 28:75]
SPACIC_alpha <- cronbach.alpha(SPACIC_items)
SPACIC_alpha

Kfull_items <- descr_data_full_before[, 28:43]
Kfull_alpha <- cronbach.alpha(Kfull_items)
Kfull_alpha

Kdir_items <- descr_data_full_before[, 28:35]
Kdir_alpha <- cronbach.alpha(Kdir_items)
Kdir_alpha

Kindir_items <- descr_data_full_before[, 36:43]
Kindir_alpha <- cronbach.alpha(Kindir_items)
Kindir_alpha

Cfull_items <- descr_data_full_before[, 44:59]
Cfull_alpha <- cronbach.alpha(Cfull_items)
Cfull_alpha

Cdir_items <- descr_data_full_before[, 44:51]
Cdir_alpha <- cronbach.alpha(Cdir_items)
Cdir_alpha

Cindir_items <- descr_data_full_before[, 52:59]
Cindir_alpha <- cronbach.alpha(Cindir_items)
Cindir_alpha

Wfull_items <- descr_data_full_before[, 60:75]
Wfull_alpha <- cronbach.alpha(Wfull_items)
Wfull_alpha

Wdir_items <- descr_data_full_before[, 60:67]
Wdir_alpha <- cronbach.alpha(Wdir_items)
Wdir_alpha

Windir_items <- descr_data_full_before[, 68:75]
Windir_alpha <- cronbach.alpha(Windir_items)
Windir_alpha
```

### RM ANOVA subset
```{r build df table 2b, include = FALSE}
descr_subs_df2b <- ddply(descr_data_subs, ~ Treatment*Timing, summarise, 
                       n_pupils = length(Full_SPACIP_mean), 
                       MeanSPACIC = mean(Full_SPACIP_mean), 
                       SESPACIC = sd(Full_SPACIP_mean)/sqrt(length(Full_SPACIP_mean)), 
                       MeanKfull = mean(K_mean), 
                       SEKfull = sd(K_mean)/sqrt(length(K_mean)),
                       MeanK_mean_dir = mean(K_mean_dir), 
                       SEK_mean_dir = sd(K_mean_dir)/sqrt(length(K_mean_dir)),
                       MeanK_mean_ind = mean(K_mean_ind), 
                       SEK_mean_ind = sd(K_mean_ind)/sqrt(length(K_mean_ind)), 
                       MeanCfull = mean(C_mean), 
                       SECfull = sd(C_mean)/sqrt(length(C_mean)),
                       MeanC_mean_dir = mean(C_mean_dir), 
                       SEC_mean_dir = sd(C_mean_dir)/sqrt(length(C_mean_dir)),
                       MeanC_mean_ind = mean(C_mean_ind), 
                       SEC_mean_ind = sd(C_mean_ind)/sqrt(length(C_mean_ind)), 
                       MeanWfull = mean(W_mean), 
                       SEWfull = sd(W_mean)/sqrt(length(W_mean)),
                       MeanW_mean_dir = mean(W_mean_dir), 
                       SEW_mean_dir = sd(W_mean_dir)/sqrt(length(W_mean_dir)),
                       MeanW_mean_ind = mean(W_mean_ind), 
                       SEW_mean_ind = sd(W_mean_ind)/sqrt(length(W_mean_ind)))
descr_subs_df2b
```

#### Cronbach's alpha
```{r calculate for the full dataset, include = FALSE}
str(descr_data_subs)

subs <- descr_data_subs[, 28:35]
str(subs)

SPACIC_items <- descr_data_full[, 28:75]
SPACIC_alpha <- cronbach.alpha(SPACIC_items)
SPACIC_alpha

Kfull_items <- descr_data_full[, 28:43]
Kfull_alpha <- cronbach.alpha(Kfull_items)
Kfull_alpha

Kdir_items <- descr_data_full[, 28:35]
Kdir_alpha <- cronbach.alpha(Kdir_items)
Kdir_alpha

Kindir_items <- descr_data_full[, 36:43]
Kindir_alpha <- cronbach.alpha(Kindir_items)
Kindir_alpha

Cfull_items <- descr_data_full[, 44:59]
Cfull_alpha <- cronbach.alpha(Cfull_items)
Cfull_alpha

Cdir_items <- descr_data_full[, 44:51]
Cdir_alpha <- cronbach.alpha(Cdir_items)
Cdir_alpha

Cindir_items <- descr_data_full[, 52:59]
Cindir_alpha <- cronbach.alpha(Cindir_items)
Cindir_alpha

Wfull_items <- descr_data_full[, 60:75]
Wfull_alpha <- cronbach.alpha(Wfull_items)
Wfull_alpha

Wdir_items <- descr_data_full[, 60:67]
Wdir_alpha <- cronbach.alpha(Wdir_items)
Wdir_alpha

Windir_items <- descr_data_full[, 68:75]
Windir_alpha <- cronbach.alpha(Windir_items)
Windir_alpha
```

## Table S1

### Full dataset
```{r build df supp table 1, include = FALSE}
descr_full_df_K <- ddply(descr_data_full, ~ Treatment*Timing, summarise, 
                       n_pupils = length(Full_SPACIP_mean), 
                       MeanK_habitats = mean(K_habitats), 
                       SEK_habitats = sd(K_habitats)/sqrt(length(K_habitats)), 
                       MeanK_gardening = mean(K_gardening), 
                       SEK_gardening = sd(K_gardening)/sqrt(length(K_gardening)),
                       MeanK_growingplants = mean(K_growingplants), 
                       SEK_growingplants = sd(K_growingplants)/sqrt(length(K_growingplants)),
                       MeanK_lawn = mean(K_lawn), 
                       SEK_lawn = sd(K_lawn)/sqrt(length(K_lawn)), 
                       MeanK_avoidpesticides = mean(K_avoidpesticides), 
                       SEK_avoidpesticides = sd(K_avoidpesticides)/sqrt(length(K_avoidpesticides)),
                       MeanK_providenests = mean(K_providenests), 
                       SEK_providenests = sd(K_providenests)/sqrt(length(K_providenests)),
                       MeanK_trackdiversity = mean(K_trackdiversity), 
                       SEK_trackdiversity = sd(K_trackdiversity)/sqrt(length(K_trackdiversity)), 
                       MeanK_schoolyard = mean(K_schoolyard), 
                       SEK_schoolyard = sd(K_schoolyard)/sqrt(length(K_schoolyard)),
                       MeanK_informhowtohelp = mean(K_informhowtohelp), 
                       SEK_informhowtohelp = sd(K_informhowtohelp)/sqrt(length(K_informhowtohelp)),
                       MeanK_raiseawareness = mean(K_raiseawareness), 
                       SEK_raiseawareness = sd(K_raiseawareness)/sqrt(length(K_raiseawareness)), 
                       MeanK_positiveattitude = mean(K_positiveattitude), 
                       SEK_positiveattitude = sd(K_positiveattitude)/sqrt(length(K_positiveattitude)), 
                       MeanK_insectfrproducts = mean(K_insectfrproducts), 
                       SEK_insectfrproducts = sd(K_insectfrproducts)/sqrt(length(K_insectfrproducts)),
                       MeanK_organisations = mean(K_organisations), 
                       SEK_organisations = sd(K_organisations)/sqrt(length(K_organisations)),
                       MeanK_contributetoresearch = mean(K_contributetoresearch), 
                       SEK_contributetoresearch = sd(K_contributetoresearch)/sqrt(length(K_contributetoresearch)),
                       MeanK_contactcompanies = mean(K_contactcompanies), 
                       SEK_contactcompanies = sd(K_contactcompanies)/sqrt(length(K_contactcompanies)), 
                       MeanK_supportactions = mean(K_supportactions), 
                       SEK_supportactions = sd(K_supportactions)/sqrt(length(K_supportactions)))
descr_full_df_K
```

### RM ANOVA subset
```{r build df table 2b, include = FALSE}
#descr_subs_df_K <- ddply(descr_data_full, ~ Treatment*Timing, summarise, 
                       n_pupils = length(Full_SPACIP_mean), 
                       MeanK_habitats = mean(K_habitats), 
                       SEK_habitats = sd(K_habitats)/sqrt(length(K_habitats)), 
                       MeanK_gardening = mean(K_gardening), 
                       SEK_gardening = sd(K_gardening)/sqrt(length(K_gardening)),
                       MeanK_growingplants = mean(K_growingplants), 
                       SEK_growingplants = sd(K_growingplants)/sqrt(length(K_growingplants)),
                       MeanK_lawn = mean(K_lawn), 
                       SEK_lawn = sd(K_lawn)/sqrt(length(K_lawn)), 
                       MeanK_avoidpesticides = mean(K_avoidpesticides), 
                       SEK_avoidpesticides = sd(K_avoidpesticides)/sqrt(length(K_avoidpesticides)),
                       MeanK_providenests = mean(K_providenests), 
                       SEK_providenests = sd(K_providenests)/sqrt(length(K_providenests)),
                       MeanK_trackdiversity = mean(K_trackdiversity), 
                       SEK_trackdiversity = sd(K_trackdiversity)/sqrt(length(K_trackdiversity)), 
                       MeanK_schoolyard = mean(K_schoolyard), 
                       SEK_schoolyard = sd(K_schoolyard)/sqrt(length(K_schoolyard)),
                       MeanK_informhowtohelp = mean(K_informhowtohelp), 
                       SEK_informhowtohelp = sd(K_informhowtohelp)/sqrt(length(K_informhowtohelp)),
                       MeanK_raiseawareness = mean(K_raiseawareness), 
                       SEK_raiseawareness = sd(K_raiseawareness)/sqrt(length(K_raiseawareness)), 
                       MeanK_positiveattitude = mean(K_positiveattitude), 
                       SEK_positiveattitude = sd(K_positiveattitude)/sqrt(length(K_positiveattitude)), 
                       MeanK_insectfrproducts = mean(K_insectfrproducts), 
                       SEK_insectfrproducts = sd(K_insectfrproducts)/sqrt(length(K_insectfrproducts)),
                       MeanK_organisations = mean(K_organisations), 
                       SEK_organisations = sd(K_organisations)/sqrt(length(K_organisations)),
                       MeanK_contributetoresearch = mean(K_contributetoresearch), 
                       SEK_contributetoresearch = sd(K_contributetoresearch)/sqrt(length(K_contributetoresearch)),
                       MeanK_contactcompanies = mean(K_contactcompanies), 
                       SEK_contactcompanies = sd(K_contactcompanies)/sqrt(length(K_contactcompanies)), 
                       MeanK_supportactions = mean(K_supportactions), 
                       SEK_supportactions = sd(K_supportactions)/sqrt(length(K_supportactions)))
```

## Table S2

### Full dataset
```{r build df supp table 2, include = FALSE}
descr_full_df_C <- ddply(descr_data_full, ~ Treatment*Timing, summarise, 
                       n_pupils = length(Full_SPACIP_mean), 
                       MeanC_habitats = mean(C_habitats), 
                       SEC_habitats = sd(C_habitats)/sqrt(length(C_habitats)), 
                       MeanC_gardening = mean(C_gardening), 
                       SEC_gardening = sd(C_gardening)/sqrt(length(C_gardening)),
                       MeanC_growingplants = mean(C_growingplants), 
                       SEC_growingplants = sd(C_growingplants)/sqrt(length(C_growingplants)),
                       MeanC_lawn = mean(C_lawn), 
                       SEC_lawn = sd(C_lawn)/sqrt(length(C_lawn)), 
                       MeanC_avoidpesticides = mean(C_avoidpesticides), 
                       SEC_avoidpesticides = sd(C_avoidpesticides)/sqrt(length(C_avoidpesticides)),
                       MeanC_providenests = mean(C_providenests), 
                       SEC_providenests = sd(C_providenests)/sqrt(length(C_providenests)),
                       MeanC_trackdiversity = mean(C_trackdiversity), 
                       SEC_trackdiversity = sd(C_trackdiversity)/sqrt(length(C_trackdiversity)), 
                       MeanC_schoolyard = mean(C_schoolyard), 
                       SEC_schoolyard = sd(C_schoolyard)/sqrt(length(C_schoolyard)),
                       MeanC_informhowtohelp = mean(C_informhowtohelp), 
                       SEC_informhowtohelp = sd(C_informhowtohelp)/sqrt(length(C_informhowtohelp)),
                       MeanC_raiseawareness = mean(C_raiseawareness), 
                       SEC_raiseawareness = sd(C_raiseawareness)/sqrt(length(C_raiseawareness)), 
                       MeanC_positiveattitude = mean(C_positiveattitude), 
                       SEC_positiveattitude = sd(C_positiveattitude)/sqrt(length(C_positiveattitude)), 
                       MeanC_insectfrproducts = mean(C_insectfrproducts), 
                       SEC_insectfrproducts = sd(C_insectfrproducts)/sqrt(length(C_insectfrproducts)),
                       MeanC_organisations = mean(C_organisations), 
                       SEC_organisations = sd(C_organisations)/sqrt(length(C_organisations)),
                       MeanC_contributetoresearch = mean(C_contributetoresearch), 
                       SEC_contributetoresearch = sd(C_contributetoresearch)/sqrt(length(C_contributetoresearch)),
                       MeanC_contactcompanies = mean(C_contactcompanies), 
                       SEC_contactcompanies = sd(C_contactcompanies)/sqrt(length(C_contactcompanies)), 
                       MeanC_supportactions = mean(C_supportactions), 
                       SEC_supportactions = sd(C_supportactions)/sqrt(length(C_supportactions)))
descr_full_df_C
```

## Table S3

### Full dataset
```{r build df supp table 3, include = FALSE}
descr_full_df_W <- ddply(descr_data_full, ~ Treatment*Timing, summarise, 
                       n_pupils = length(Full_SPACIP_mean), 
                       MeanW_habitats = mean(W_habitats), 
                       SEW_habitats = sd(W_habitats)/sqrt(length(W_habitats)), 
                       MeanW_gardening = mean(W_gardening), 
                       SEW_gardening = sd(W_gardening)/sqrt(length(W_gardening)),
                       MeanW_growingplants = mean(W_growingplants), 
                       SEW_growingplants = sd(W_growingplants)/sqrt(length(W_growingplants)),
                       MeanW_lawn = mean(W_lawn), 
                       SEW_lawn = sd(W_lawn)/sqrt(length(W_lawn)), 
                       MeanW_avoidpesticides = mean(W_avoidpesticides), 
                       SEW_avoidpesticides = sd(W_avoidpesticides)/sqrt(length(W_avoidpesticides)),
                       MeanW_providenests = mean(W_providenests), 
                       SEW_providenests = sd(W_providenests)/sqrt(length(W_providenests)),
                       MeanW_trackdiversity = mean(W_trackdiversity), 
                       SEW_trackdiversity = sd(W_trackdiversity)/sqrt(length(W_trackdiversity)), 
                       MeanW_schoolyard = mean(W_schoolyard), 
                       SEW_schoolyard = sd(W_schoolyard)/sqrt(length(W_schoolyard)),
                       MeanW_informhowtohelp = mean(W_informhowtohelp), 
                       SEW_informhowtohelp = sd(W_informhowtohelp)/sqrt(length(W_informhowtohelp)),
                       MeanW_raiseawareness = mean(W_raiseawareness), 
                       SEW_raiseawareness = sd(W_raiseawareness)/sqrt(length(W_raiseawareness)), 
                       MeanW_positiveattitude = mean(W_positiveattitudes), 
                       SEW_positiveattitude = sd(W_positiveattitudes)/sqrt(length(W_positiveattitudes)), 
                       MeanW_insectfrproducts = mean(W_insectfrproducts), 
                       SEW_insectfrproducts = sd(W_insectfrproducts)/sqrt(length(W_insectfrproducts)),
                       MeanW_organisations = mean(W_organisations), 
                       SEW_organisations = sd(W_organisations)/sqrt(length(W_organisations)),
                       MeanW_contributetoresearch = mean(W_contributetoresearch), 
                       SEW_contributetoresearch = sd(W_contributetoresearch)/sqrt(length(W_contributetoresearch)),
                       MeanW_contactcompanies = mean(W_contactcompanies), 
                       SEW_contactcompanies = sd(W_contactcompanies)/sqrt(length(W_contactcompanies)), 
                       MeanW_supportactions = mean(W_supportaction), 
                       SEW_supportactions = sd(W_supportaction)/sqrt(length(W_supportaction)))
descr_full_df_W
```

## Table S4

### Full dataset
```{r build df supp table 4a, include = FALSE}
schools_full_df <- ddply(descr_data_full, ~ Treatment*Timing*School, summarise, 
                       n_pupils = length(Full_SPACIP_mean))
schools_full_df
```

### RM ANOVA subset
```{r build df supp table 4b, include = FALSE}
schools_subs_df <- ddply(descr_data_subs, ~ Treatment*Timing*School, summarise, 
                       n_pupils = length(Full_SPACIP_mean))
schools_subs_df
```